\documentclass[12pt, letterpaper]{book}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{setspace}
\usepackage{parskip}

% Page geometry
\geometry{
    letterpaper,
    margin=1.25in,
    top=1.5in,
    bottom=1.5in
}

% Fix fancyhdr headheight warning
\setlength{\headheight}{14.49998pt}

% Chapter formatting for poems
\titleformat{\chapter}[display]
    {\normalfont\huge\bfseries\centering}
    {}
    {0pt}
    {\Huge}
\titlespacing*{\chapter}{0pt}{-30pt}{40pt}

% Header and footer
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[RE]{\textit{Faiyaz's Poems}}
\fancyhead[LO]{\textit{\leftmark}}
\renewcommand{\headrulewidth}{0.4pt}

% Line spacing
\onehalfspacing

% Title information
\title{\Huge\textbf{Faiyaz's Poems}}
\author{\Large Faiyaz Chowdhury}
\date{}

\begin{document}

% Title page
\begin{titlepage}
    \centering
    \vspace*{2in}
    {\Huge\bfseries Faiyaz's Poems\par}
    \vspace{1in}
    {\Large Faiyaz Chowdhury\par}
    \vfill
\end{titlepage}

% Table of contents
\tableofcontents
\newpage

% The Poems

\chapter{Water and Stone}

\begin{verse}
Water and Stone start at the mountain top\\
One formed from wet cloud in sky,\\
The other, from hot earth as rock.
\end{verse}

\begin{verse}
Water, with full force and might,\\
Flows anywhere, but with no end in sight.\\
Stone with all form and structure,\\
Forms the course, but no way to get there
\end{verse}

\begin{verse}
Water flows down the track, crashing with Stone,\\
Havoc on impact, a tempest on land.\\
But through time both are shown,\\
A peaceful soft white sand.
\end{verse}

\begin{verse}
The crashes continue down the mountain track,\\
Water disturbs Stone, while Stone blocks the path.\\
As the years and years pass, water pulls Stone, while Stone shows the way,\\
What started as conflict had past, Becoming of love, help and play
\end{verse}

\begin{verse}
At last, they reach the coast and basin,\\
A great deal away from fight and sin.\\
Though they think they arrived at paradise\\
They actually made it through their own sacrifice.
\end{verse}

\begin{verse}
The sand at bottom, was born from Water and Stone at top,\\
Where the Water and Stone had learnt together, to let the fight stop.\\
Once they reach the bottom of terrain,\\
The cycle resets, to being their journey again.
\end{verse}


\chapter{Real}

\begin{verse}
The world is full of death, war, decay,\\
And so we ask, why must it be this way?\\
We watch as our hearts feel dismay,\\
And slowly, our ideals, we often may betray.
\end{verse}

\begin{verse}
However, I say, these things they are not real.\\
And Apple is red, I can touch taste and feel.
\end{verse}

\begin{verse}
Beware that I ask what an Apple is not.\\
For negative space is infinite, a thought laden with distraught.\\
An Apple is not a smile, nor a hot cross bun.\\
Not love, the sky, a song, the moon or the sun.
\end{verse}

\begin{verse}
As such, death is not real, but the absence of life.\\
With expectation out of the way, we can free our hearts of strife.\\
And life is real, indeed the greatest gift and joy.\\
And with time, with this positive shift, something we can enjoy.
\end{verse}


\chapter{Love}

\begin{verse}
There is a special thing called love,\\
Something that's always just enough,\\
Someone who makes you feel at home,\\
No matter how close or far you roam.
\end{verse}

\begin{verse}
May now that love travel with you,\\
To places and goals you may pursue,\\
Know now that my concern is true,\\
Your feelings and heart are sacred too.
\end{verse}


\end{document}

